---
title: "R markdown"
author: "Elizabeth Josephine"
date: "2023-07-25"
output:
  html_document:
    df_print: paged
---

```{r}
## 
```


```{r}
load.libraries <- c('data.table', 'testthat', 'gridExtra', 'corrplot', 'GGally', 'ggplot2', 'e1071', 'dplyr')
install.lib <- load.libraries[!load.libraries %in% installed.packages()]
for(libs in install.lib) install.packages(libs, dependences = TRUE)
sapply(load.libraries, require, character = TRUE)
```

```{r}
#reading data
df <- read.csv("deworming_children.csv")
```

# DATA CHECKING
```{r}
#data preview
View(df)
```

```{r}
class(df)
```


```{r}
# data head preview
head(df)
```

```{r}
# data tail preview
tail(df)
```

```{r}
#column names
colnames(df)
```

```{r}
#checking datatypes
sapply(df, class)
```

```{r}
#data structure
str(df)
```

```{r}
#data dimension
dim(df) # there are 1410 rows and 11 columns
```

# DATA CLEANING
## Missing Values
```{r}
# checking for missing values
sum(is.na(df)) #there are 399 missing values in the data
```

```{r}
# dropping rows with missing values
df1 <- na.omit(df)
View(df1)
df1
```

```{r}
# printing df1
write.csv(df1, file = "clean_deworm.csv", row.names = FALSE)
```

```{r}
#displaying all rows from the dataset which don't contain any missing values 

dim(df1) # there are 1048 rows and 11 columns
```

## Duplicates
```{r}
# checking for duplicates in new dataset
duplicated_rows <- df1[duplicated(df1),]
duplicated_rows # there are no duplicates in the data
```

```{r}
# selecting the numerical data columns
df2 <- df1 %>% select_if(is.numeric)
View(df2)
df2
```

```{r}
#viewing the remaining columns
colnames(df2)
```
## Checking for outliers
```{r}
# visualizing any existing outliers using a boxplot
boxplot(df2) #there are outliers in the data. However, we will keep the outliers a s they are essential to the analysis
```
*#### removing outliers using z-score*

remove_outliers_zscore <- function(df2, threshold) {
  z_scores <- scale(df2)  # Calculate Z-scores for each column in the dataframe
  outlier_rows <- apply(z_scores, 1, function(row) any(abs(row) > threshold))
  cleaned_data <- df2[!outlier_rows, ]
  
  return(cleaned_data)
}

data_cleaned <- remove_outliers_zscore(df2, threshold = 3)
****

# EXPLORATORY DATA ANALYSIS
# FEATURE ENGINEERING

```{r}
#install.packages("gapminder")
library(gapminder)
```

```{r}
df3<- select(df1, period, county, Total.Dewormed)
View(df3)
```

```{r}
df4<- df3 %>% pivot_wider(names_from = period, values_from = Total.Dewormed)
View(df4)
dim(df4)
```

```{r}
# checking for missing values
sum(is.na(df4)) #there are 399 missing values in the data
```

```{r}
# dropping rows with missing values
df5 <- na.omit(df4)
View(df5)
df5
dim(df5)
```

```{r}
# printing df1
write.csv(df5, file = "clean_deworm4.csv", row.names = FALSE)
```

```{r}
#displaying all rows from the dataset which don't contain any missing values 

dim(df5) # there are 1048 rows and 11 columns
```

## Duplicates
```{r}
# checking for duplicates in new dataset
duplicated_rows <- df5[duplicated(df5),]
duplicated_rows # there are no duplicates in the data
```

## UNIVARITE ANALYSIS
```{r}
summary(df5)
```

```{r}
#descriptive statistics of the numerical columns
library(Hmisc)
describe(df5)
```

```{r}
# selecting the numerical data columns
df6 <- df5 %>% select_if(is.numeric)
View(df6)
df6
```
### Univariate Graphical

```{r}
# creating a boxplot graph for the variable all the numerical variables
boxplot(df6)
```

## BIVARIATE ANALYSIS
### covariance
```{r}
# covariance of variables
cov <- cov(df6)
round(cov, 2)
```

### correlation
```{r}
# finding the correlation matrix
library(corrplot)
```

```{r}
coo <- cor(df6)
round(coo, 2)
```

# implementation of the solution
## principal component analysis
```{r}
df4 <- prcomp(df3)
summary(df4)
```

```{r}
# Calling str() to have a look at your PCA object
#str(df4)
```

```{r}
# creating a scatterplot
#plot(Age, Daily_Internet_Usage, xlab="Age", ylab="Daily_Internet_Usage")
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

